---
title: "Population Dynamics"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Population Dynamics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.asp = 0.618
)

library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(Silvanus)
library(furrr)
```

# Introduction

Get the life table here and talk about the basic 

First we'll create a test population of 15 25-year-olds.

```{r}
pop1 <- create_individuals(15)
```


# Unbounded growth
Then run a 500 year simulation from that initial population.

```{r}
set.seed(1000) # set seed for reproducibility
nsim <- 500

demographic_sim <- tibble(time = 0:nsim,
                          age = accumulate(1:nsim, ~population_dynamics(.x), .init = pop1),
                          population = map_dbl(age, length)) %>%
  mutate(dependency_ratio = map_dbl(age,  ~sum(!between(.x, 15, 65)) / sum(between(.x, 15, 65))) * 100)
```

Look at the results.

## Dependency ratio

```{r, fig.width = 6, fig.asp = 1, echo = FALSE}
a <-   ggplot(demographic_sim, aes(time, dependency_ratio))+
  geom_line() +
  labs(x = 'Year', y = 'Dependency ratio') +
  theme_bw()
d1 <- demographic_sim %>%
  filter(time %in% c(0, 100, 200, 300, 400, 500)) %>%
  unnest(cols = age) %>%
  mutate(time = paste('Year', time)) %>%
  group_by(time) %>%
  mutate(median_age = median(age))
b <- ggplot(d1, aes(age)) +
  geom_histogram(binwidth = 1, center = 0) +
  geom_vline(aes(xintercept = median_age), color = 'red', linetype = 2) + # find a way to get median age to vary with facet
  facet_wrap(~time, scales = 'free_y') +
  labs(x = 'Age', y = 'Count') +
 # labs(title = paste('Evolution of the age distribution over', nsim, 'years'), subtitle = 'Century snapshots') +
  theme_bw()

a / b + plot_annotation(tag_levels = 'A')
```

## Growth rates


```{r, cache = TRUE}
plan(multisession)

r_test <- future_map(1:100, ~ 
                       tibble(time = 0:600,
                          age = accumulate(1:600, ~population_dynamics(.x), .init = pop1),
                          population = map_dbl(age, length)) %>%
                       select(-age), .progress = TRUE) %>%
  bind_rows(.id = 'rep')
```


Fit a poisson GLM with a log link function and a varying intercept to represent the different effective initial conditions of each simulation run.
```{r}
m1 <- glm(population ~ time + rep, data = r_test, family = poisson(link = 'log'))
```

From the GLM, we see that the intrinsic growth rate is `r round(coefficients(m1)['time'] * 100, 2)`% per year.


```{r, echo = FALSE}
ggplot(r_test, aes(time, population)) +
  geom_line(aes( group = rep), alpha = .1) +
  geom_smooth(method = 'lm', color = 'red') +
  scale_y_log10() +
  theme_classic()
```


# Food-limited growth

Now we can look at a food-limited population.

```{r}
plan(multisession)
demographic_sim2  <- expand_grid(rep = 1:25, 
                                 food_ratio = seq(0.7, 1, by = 0.05)) %>%
  mutate(sim = future_map(food_ratio, function(x) accumulate(1:nsim, ~population_dynamics(.x, food_ratio_c = x), .init = pop1), .progress = TRUE),
         sim = map(sim, ~tibble(time = 0:nsim,
                          age = .,
                          population = map_dbl(age, length)) %>%
                     mutate(dependency_ratio = map_dbl(age,  ~sum(!between(.x, 15, 65))/sum(between(.x, 15, 65))) * 100)))

```


```{r}
demographic_sim2 %>% unnest %>%  filter(time == nsim) %>%
  ggplot(aes(food_ratio, population, group = food_ratio)) +
  geom_boxplot() +
  theme_classic()

demographic_sim2 %>% unnest %>%  filter(time == nsim) %>%
  ggplot(aes(food_ratio, dependency_ratio, group = food_ratio)) +
  geom_boxplot() +
  theme_classic()
```

```{r}
demographic_sim2 %>%
  unnest %>%
 # filter(food_ratio >= .8) %>%
  ggplot(aes(time, population, group =interaction(rep, food_ratio), color = food_ratio)) +
  geom_line(alpha = .5) +
  #scale_y_continuous(trans = 'log') +
  scale_color_viridis_c() +
  theme_bw()

demographic_sim2 %>%
  unnest %>%
 filter(food_ratio >= .8) %>%
  ggplot(aes(time, population, group =interaction(rep, food_ratio), color = food_ratio)) +
  geom_line(alpha = .5) +
  scale_y_continuous(trans = 'log') +
  scale_color_viridis_c() +
  theme_bw()
```


## Drought response

```{r}
nsim <- 500
rn <- c(rep(1, 300), rep(0.5, 30), rep(1, 170))
demographic_sim3 <- tibble(time = 0:nsim,
                          age = accumulate(rn, ~population_dynamics(.x, food_ratio_c = .y), .init = pop1),
                          population = map_dbl(age, length)) %>%
  mutate(dependency_ratio = map_dbl(age,  ~sum(!between(.x, 15, 65))/sum(between(.x, 15, 65))) * 100)
```

```{r, echo = FALSE}
ggplot(demographic_sim3, aes(time, population)) + 
  geom_ribbon(aes(xmin = 300, xmax = 330), alpha = 0.2) +
  geom_line() +
  theme_classic()


ggplot(demographic_sim3, aes(time, dependency_ratio)) + 
  geom_ribbon(aes(xmin = 300, xmax = 330), alpha = 0.2) +
  geom_line() +
  theme_classic()
```


